```{r}
library(tidyverse)
library(tidycensus)
library(haven)

Food_data <-read_xpt("C:/Users/user/Downloads/P_DR1IFF.xpt")
demographics<-read_xpt("C:/Users/user/Downloads/P_DEMO.xpt")

summary(Food_data)

```

```{r}
#Lets remove the unnecessary variables from Food data 


Food_data <- select(Food_data, SEQN, DR1ILINE, DRDINT, DR1_040Z, DR1IFDCD, DR1IGRMS, DR1ISUGR)

##lets change the variable names to proper format

colnames(Food_data)[2] <- "Food_number"
colnames(Food_data)[3] <- "Number_of_days_intake"
colnames(Food_data)[4] <- "Did_eat_the_food_at-home"
colnames(Food_data)[5] <- "food_code"
colnames(Food_data)[6] <- "Grams"
colnames(Food_data)[7] <- "Sugar"
#========================================================================

#Lets remove the unnecessary variables from demographics 

demographics <- select(demographics, SEQN, RIAGENDR, RIDAGEYR,DMDBORN4, RIDRETH3,DMDEDUC2, DMDMARTZ, INDFMPIR)

#lets change the variable names to proper format

colnames(demographics)[2] <- "MALE"
colnames(demographics)[3] <- "AGE"
colnames(demographics)[4] <- "Country of Birth"
colnames(demographics)[5] <- "Race"
colnames(demographics)[6] <- "EDUCATION"
colnames(demographics)[7] <- "MARITAL"
colnames(demographics)[8] <- "Family_income_to_poverty"

# Now lets filter age by 20+ (We dont need food preferences of teenagers, Also it will remove a lot of NAs that we have in Education level, some foods, martial status, language of interview and other variables)

demographics <- demographics %>% filter(AGE>= 20)

# also to better match the ACS data we will remove the 2 categories from education level as they are only 15 observations
demographics<-demographics[demographics$`EDUCATION`!= "7",]
demographics<-demographics[demographics$`EDUCATION`!= "9",]

#==================================================================

# lets merge food data with demographic data

merged_set<- merge(Food_data, demographics , by = "SEQN")

```

# Now lets mege the food code with the food description(names)

```{r}

library(dplyr)

library("readxl")

foodnames<- read_excel("C:/Users/user/Downloads/WWEIA2017_March2020_foodcat_FNDDS.xlsx")

# merging data by food code

merged_set_final <- merge(x=merged_set,y=foodnames, 
          by="food_code", all.x=TRUE)

merged_set_final<- merged_set_final %>% relocate(food_code, .before=food_code_description)


```

# Now to mege it with ACS PUMS data

## Recode demographic variables according to American Community Survey standards

Export PUMS variable definitions.

```{r}
library(tidycensus)
library(tidyverse)
library(rio)
pums_vars_2020 <- pums_variables%>%
  filter(survey=="acs5",
         year==2020,
         level == "person")

export(pums_vars_2020,"pums_vars_2021.xlsx")

```

Get PUMS data ACS data contains 386061 respondents representing the whole population of California.

```{r}

gc()
memory.limit(9999999999)
gc() 

Western_states <- c("WA", "OR", "ID", "NV", "CA")
Western_states1 <- c( "UT", "AZ", "NM", "TX")

pums1 <- get_pums(
  variables = c("PUMA", 
                "PWGTP",
                "SEX", 
                "AGEP", 
                "SCHL",
                "HISP",
                "RACAIAN",
                "RACASN",
                "RACBLK",
                "RACNH",
                "RACPI",
                "RACSOR",
                "RACWHT",
                "RACNUM",
                "MAR",
                "POVPIP",
                "CIT"),
  state = Western_states1,
  survey = "acs5",
  year = 2020)

Pums2 <- rbind(pums, pums1)
pums <- Pums2

export(pums,"pums_2020.xlsx")

```

#Recode PUMS variables

```{r}
 
pums_final<-pums%>%transmute(
  PUMA=PUMA,
  PWGTP=PWGTP,
  AGE=AGEP,
  POVPIP= POVPIP/100,
  EDUCATION=case_when(
    SCHL %in% c("01","02","03","04","05","06","07",
                "08","09","10","11")~"Less than high school degree",
    SCHL %in% c("12","13","14","15")~ "9-11th grade_Includes 12th grade with no diploma",
    SCHL %in% c("16","17")~ "High school graduate/GED or equivalent",
    SCHL %in% c("18","19","20")~ "Some college or AA degree",
    SCHL %in% c("21","22","23","24")~ "College graduate or above", 
    TRUE~"Unknown"),
  MALE=as.numeric(SEX=="1"),
  Race=case_when(
    HISP %in% c("02")~"Mexican American",
    HISP %in% c("03","04","05","06","07","08","09","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24")  ~"Other Hispanic",
  RACWHT %in% c("1")~"White",
  RACBLK %in% c("1")~"Black",
  RACASN %in% c("1")~"Asian",
  RACNUM %in% c("1","2","03","04","05","06")~"Other Race - Including Multi-Racial"),
    RACE_NAT_HAW=as.numeric(RACNH),
  RACE_PAC_ISL=as.numeric(RACPI),
  RACE_NAT_AM=as.numeric(RACAIAN),
MARITAL=case_when(
MAR=="1"~"Married",
MAR%in% c("2","3","4")~"Separated, divorced or widowed",
MAR=="5"~"Single"),
CIT=case_when(
    CIT=="1"~"1",
    TRUE~"2")
)

# lets convert income to poverty ratio to the same values as of Food data
pums_final <- mutate(pums_final, POVPIP = case_when(
  POVPIP == -0.01 ~ 0, POVPIP == 5.01~5,
  TRUE   ~ POVPIP 
))

# Also since we filterd age in survey data lets filter age by 20 in census data too

pums_final <- pums_final %>% filter(AGE >= 20)

#Also

pums_final$Gender <- factor(pums_final$MALE, levels = c(0, 1), labels = c("Female", "Male"))


pums_final %>% count(`Race`, sort = TRUE)

```

## Convert character variables to factors

```{r}

pums_final$EDUCATION<- as.factor(pums_final$EDUCATION)
pums_final$Race<- as.factor(pums_final$Race)
pums_final$MARITAL<- as.factor(pums_final$MARITAL)
pums_final$CIT<- as.factor(pums_final$CIT)
pums_final$Gender<- as.factor(pums_final$Gender)

#rename columns
colnames(pums_final)[colnames(pums_final) == "CIT"] ="Country of Birth"
colnames(pums_final)[colnames(pums_final) == "POVPIP"] ="Family_income_to_poverty"

#survy data final

#change to factors Survey variables

merged_set_final$Race<- factor(merged_set_final$Race, levels = c(1, 2, 3,4,6,7), labels = c("Mexican American", "Other Hispanic","White", "Black", "Asian", "Other Race - Including Multi-Racial"))

merged_set_final$EDUCATION<- factor(merged_set_final$EDUCATION, levels = c(1, 2, 3,4,5), labels = c("Less than high school degree", "9-11th grade_Includes 12th grade with no diploma", "High school graduate/GED or equivalent", "Some college or AA degree", "College graduate or above"))

merged_set_final$Gender <- factor(merged_set_final$MALE, levels = c(1, 2), labels = c("Male", "Female"))
merged_set_final$MARITAL <- factor(merged_set_final$MARITAL, levels = c(1, 2,3), labels = c("Married", "Separated, divorced or widowed", "Single"))

merged_set_final<-merged_set_final[merged_set_final$`Did_eat_the_food_at-home` != "9",]

merged_set_final$`Did_eat_the_food_at-home` <- factor(merged_set_final$`Did_eat_the_food_at-home`, levels = c(1, 2), labels = c("Yes", "No"))

merged_set_final$Gender <- as.factor(merged_set_final$Gender)
merged_set_final$`Did_eat_the_food_at-home` <- as.factor(merged_set_final$`Did_eat_the_food_at-home`)
merged_set_final$`Country of Birth` <- as.factor(merged_set_final$`Country of Birth`)
merged_set_final$EDUCATION <- as.factor(merged_set_final$EDUCATION)
merged_set_final$`MARITAL` <- as.factor(merged_set_final$`MARITAL`)
merged_set_final$food_code_description <- as.factor(merged_set_final$food_code_description)
merged_set_final$category_description <- as.factor(merged_set_final$category_description) 
merged_set_final$final_category <- as.factor(merged_set_final$final_category)

# lets change the set name

survey_final<- merged_set_final

```

## Remove non-existent race

There are no Pacific islanders in the survey, so we remove this variable.

```{r}
pums_final<-pums_final%>%
  select(-any_of("RACE_PAC_ISL"))

pums_final<-pums_final%>%
  select(-any_of("RACE_NAT_HAW"))

pums_final<-pums_final%>%
  select(-any_of("RACE_NAT_AM"))

pums_final<-pums_final%>%
  select(-any_of("Family_income_to_poverty"))

pums_final<-pums_final%>%
  select(-any_of("SEQN"))

survey_final<-survey_final%>% select(-c(reports_day1, reports_day2 ,category_number , Food_number, food_code, category_description, Number_of_days_intake, Family_income_to_poverty, SEQN, MALE, food_code_description))

```

Compare two datasets

```{r}
janitor::compare_df_cols(pums_final,survey_final)
```

```{r}
summary(survey_final)
 summary(pums_final)

survey_final<-survey_final%>%filter(!is.na(MARITAL))
summary(survey_final$MARITAL)

```

## Predictive Models of SCORE1-SCORE7

```{r}


```

## Create cluster membership for survey data

```{r}

survey_meat <-survey_final %>% filter( final_category == "Meat")
survey_Milk_Dairy <-survey_final %>% filter( final_category == "Milk&Dairy")
survey_Fast_food <-survey_final %>% filter( final_category == "Fast food")
survey_Fruits <-survey_final %>% filter( final_category == "Fruits")
survey_Vegetables <-survey_final %>% filter( final_category == "Vegetables")

```

## Multi-class regression model

```{r}

library(tidymodels)
library(ranger)
set.seed(123)


rf_spec <- rand_forest(trees=1000,
                       mtry = tune(),
  min_n = tune()) %>%
  set_engine("ranger",importance="permutation") %>%
  set_mode("regression")

rf_recipe<-recipe(Grams ~., survey_meat %>%select(Grams, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf <- workflow() %>%
  add_recipe(rf_recipe) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds <- vfold_cv(survey_meat, v = 10,strata = Grams)

tune_res <- tune_grid(
  tune_wf,
  resamples = folds,
  grid = 5)

tune_res

# select the best hyperparameters
best_auc <- select_best(tune_res, "rmse")


# finalize workflow
final_wf <- tune_wf%>%
  finalize_workflow(best_auc)

# final fit
rf_fit <- final_wf%>%
  parsnip::fit(survey_meat)

library(vip)
vip(rf_fit%>%extract_fit_engine())+theme_bw()

library(DALEXtra)

vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train <- 
  survey_final %>% 
  select(all_of(vip_features))


explainer_rf <- 
  explain_tidymodels(
    rf_fit, 
    data = vip_train, 
    y = survey_meat$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex <- survey_meat[120,]

rf_breakdown <- predict_parts(explainer = explainer_rf, new_observation = duplex)

####
library(forcats)

big_house <- vip_train[1269,]

set.seed(1802)
shap_house <- 
  predict_parts(
    explainer = explainer_rf, 
    new_observation = big_house, 
    type = "shap",
    B = 20
  )

shap_house %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###
vip_rf <- model_parts(explainer_rf, loss_function = loss_root_mean_square)

ggplot_imp(vip_rf)

####

set.seed(1805)
pdp_age <- model_profile(explainer_rf, N = 10000, variables = "Gender")

ggplot_pdp(pdp_age, Gender)  +
  labs(x = "Age", 
       y = "Grams", 
       color = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_y_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Gross living area", 
       y = "Sale Price", 
       color = NULL)

```

## model for dairy
```{r}
rf_recipe1<-recipe(Grams ~., survey_Milk_Dairy %>%select(Grams, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf1 <- workflow() %>%
  add_recipe(rf_recipe1) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds1 <- vfold_cv(survey_Milk_Dairy, v = 10,strata = Grams)

tune_res1 <- tune_grid(
  tune_wf1,
  resamples = folds1,
  grid = 5)

tune_res1

# select the best hyperparameters
best_auc1 <- select_best(tune_res1, "rmse")


# finalize workflow
final_wf1 <- tune_wf1%>%
  finalize_workflow(best_auc1)

# final fit
rf_fit1 <- final_wf1%>%
  parsnip::fit(survey_Milk_Dairy)

library(vip)
vip(rf_fit1%>%extract_fit_engine())+theme_bw()

```
## Model for Fast food
```{r}
rf_recipe2 <-recipe(Grams ~., survey_Fast_food %>%select(Grams, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf2 <- workflow() %>%
  add_recipe(rf_recipe2) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds2 <- vfold_cv(survey_Fast_food, v = 10,strata = Grams)

tune_res2 <- tune_grid(
  tune_wf2,
  resamples = folds2,
  grid = 5)

tune_res2

# select the best hyperparameters
best_auc2 <- select_best(tune_res2, "rmse")


# finalize workflow
final_wf2 <- tune_wf2%>%
  finalize_workflow(best_auc2)

# final fit
rf_fit2 <- final_wf2%>%
  parsnip::fit(survey_Fast_food)

library(vip)
vip(rf_fit2%>%extract_fit_engine())+theme_bw()

```

## model for Fruits

```{r}
rf_recipe3<-recipe(Grams ~., survey_Fruits %>%select(Grams, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf3 <- workflow() %>%
  add_recipe(rf_recipe3) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds3 <- vfold_cv(survey_Fruits, v = 10,strata = Grams)

tune_res3 <- tune_grid(
  tune_wf3,
  resamples = folds3,
  grid = 5)

tune_res3

# select the best hyperparameters
best_auc3 <- select_best(tune_res3, "rmse")


# finalize workflow
final_wf3 <- tune_wf3%>%
  finalize_workflow(best_auc3)

# final fit
rf_fit3 <- final_wf3%>%
  parsnip::fit(survey_Fruits)

library(vip)
vip(rf_fit3%>%extract_fit_engine())+theme_bw()

```

## model for Vegetable

```{r}
rf_recipe4<-recipe(Grams ~., survey_Vegetables %>%select(Grams, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf4 <- workflow() %>%
  add_recipe(rf_recipe4) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds4 <- vfold_cv(survey_Vegetables, v = 10,strata = Grams)

tune_res4 <- tune_grid(
  tune_wf4,
  resamples = folds4,
  grid = 5)

tune_res4

# select the best hyperparameters
best_auc4 <- select_best(tune_res4, "rmse")


# finalize workflow
final_wf4 <- tune_wf4%>%
  finalize_workflow(best_auc4)

# final fit
rf_fit4 <- final_wf4%>%
  parsnip::fit(survey_Vegetables)

library(vip)
vip(rf_fit4%>%extract_fit_engine())+theme_bw()

```

## model for Sugar

```{r}

rf_recipe5<-recipe(Sugar ~., survey_final %>%select(Sugar, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf5 <- workflow() %>%
  add_recipe(rf_recipe5) %>%
  add_model(rf_spec)

# 10-fold crossvalidation
set.seed(345)
folds5 <- vfold_cv(survey_final, v = 10,strata = Grams)

tune_res5 <- tune_grid(
  tune_wf5,
  resamples = folds5,
  grid = 5)

tune_res5

# select the best hyperparameters
best_auc5 <- select_best(tune_res5, "rmse")


# finalize workflow
final_wf5 <- tune_wf5%>%
  finalize_workflow(best_auc5)

# final fit
rf_fit5 <- final_wf5%>%
  parsnip::fit(survey_final)

library(vip)
vip(rf_fit5%>%extract_fit_engine())+theme_bw()

```

## where did you eat the food
```{r}

rf_spec1 <- rand_forest(trees=1000,
                       mtry = tune(),
  min_n = tune()) %>%
  set_engine("ranger",importance="permutation") %>%
  set_mode("classification")

colnames(survey_final)[1] = "Did_eat_the_food_at_home"

rf_recipe6<-recipe( Did_eat_the_food_at_home ~., survey_final %>%select(Did_eat_the_food_at_home, AGE, Gender, EDUCATION, MARITAL,Race, 'Country of Birth' ))

tune_wf6 <- workflow() %>%
  add_recipe(rf_recipe6) %>%
  add_model(rf_spec1)

# 10-fold crossvalidation
set.seed(345)
folds6 <- vfold_cv(survey_final, v = 10,strata = Grams)

tune_res6 <- tune_grid(
  tune_wf6,
  resamples = folds6,
  grid = 5)

tune_res6

# select the best hyperparameters
best_auc6 <- select_best(tune_res6, "roc_auc")


# finalize workflow
final_wf6 <- tune_wf6%>%
  finalize_workflow(best_auc6)

# final fit
rf_fit6 <- final_wf6%>%
  parsnip::fit(survey_final)

library(vip)
vip(rf_fit6%>%extract_fit_engine())+theme_bw()

```

## Predicted probabilities of each segment

```{r}

# predicted probabilities

segment_pred1<-predict(rf_fit,survey_meat,type="numeric")

survey_meat$segment_pred1 <- as.factor(segment_pred1$.pred)
 summary(segment_pred1)
 
```

```{r}

pums_final<-pums_final%>%filter(!is.na(Race))

pums_final$Meat<-predict(rf_fit,pums_final,type="numeric")$.pred
pums_final$Milk_dairy<-predict(rf_fit1,pums_final,type="numeric")$.pred
pums_final$Fast_food<-predict(rf_fit2,pums_final,type="numeric")$.pred
pums_final$Fruits<-predict(rf_fit3,pums_final,type="numeric")$.pred
pums_final$vegetables<-predict(rf_fit4,pums_final,type="numeric")$.pred
pums_final$Sugar<-predict(rf_fit5,pums_final,type="numeric")$.pred
pums_final$Eat_outside<-predict(rf_fit6,pums_final,type="class")$.pred_class

```

## Performance assessment: ROC AUC

```{r}
survey_final<-survey_final%>%filter(!is.na(MARITAL))

survey_meat$segment1<-predict(rf_fit,survey_meat,type="numeric")$.pred

# pperformance assessment
survey_final%>%
  rmse(Sugar, segment1)


# R-squared
actual <- survey_meat$Grams
R2 <- (sum((actual-survey_meat$segment1)^2)/sum((actual-mean(actual))^2))

```

```{r}

survey_final$Eat_Home_outside_pred<-predict(rf_fit6,survey_final,type="class")$.pred_class

survey_final%>%
  roc_auc(Did_eat_the_food_at_home, Eat_Home_outside_pred)


survey_final$segment_predicted<-predict(rf_fit6,survey_final)$.pred_class


auccc<-accuracy(survey_final, Did_eat_the_food_at_home  ,segment_predicted )

plot(auccc)


```

## Tabulate by PUMA

```{r}
puma_data<-pums_final%>%
  group_by(PUMA)%>%
  summarise(segment1_perc=weighted.mean(segment1,PWGTP)*100,
            pop20=sum(PWGTP),
            segment1_abs=round(pop20*segment1_perc/100,0))

dd<-log(puma_data$Sugar_perc)
                       
puma_data<-pums_final%>%
  group_by(PUMA)%>%
  summarise( pop20=sum(PWGTP),
    Meat_perc=(sum(PWGTP[Meat]) / pop20)*100,
    Milk_dairy_perc=(sum(PWGTP[Milk_dairy]) / pop20)*100,
    Fast_food_perc=(sum(PWGTP[Fast_food]) / pop20)*100,
    Fruits_perc=(sum(PWGTP[Fruits]) / pop20)*100,
    vegetables_perc=(sum(PWGTP[vegetables]) / pop20)*100,
    Sugar_perc=(sum(PWGTP[Sugar]) / pop20)*100,
    Eat_outside_perc=(sum(PWGTP[Eat_outside]) / pop20)*100,
            Sugar_abs=round(pop20*Sugar_perc/100,0),
            Eat_Home_outside_abs=round(pop20*Eat_outside_perc/100,0))

Column_100_ <- c("Meat_perc", "Milk_dairy_perc", "Fast_food_perc", "Fruits_perc", "vegetables_perc", "Sugar_perc", "Eat_outside_perc")

puma_data2 <- puma_data[paste0(Column_100_, "_scaled")] <- lapply(puma_data[Column_100_], function(x) (x/max(x))*100)


puma_data2 <- puma_data %>% mutate_at(vars(Column_100_), ~ scales::rescale(.)* 100)

export(puma_data,"puma_data2.xlsx",overwrite=TRUE)
```

```{r}
---
title: "Segment  Prevalence by Area (% of Population 20+)"
output:
  html_document:
    df_print: paged
---



library(mapview)
library(tigris)
library(sf)
library(rio)
library(tidyverse)
library(leafpop)
# PUMA boundaries
puma_boundaries <-pumas(state = "CA",year = 2020 )
puma_boundaries1 <-pumas(state = "WA",year = 2020)
puma_boundaries2 <-pumas(state = "OR",year = 2020)
puma_boundaries3 <-pumas(state = "NV",year = 2020)
puma_boundaries4 <-pumas(state = "UT",year = 2020)
puma_boundaries5 <-pumas(state = "AZ",year = 2020)
puma_boundaries6 <-pumas(state = "NM",year = 2020)  
puma_boundaries7 <-pumas(state = "TX",year = 2020)
puma_boundaries8 <-pumas(state = "ID",year = 2020)

puma_boundaries0 <- rbind(puma_boundaries, puma_boundaries2, puma_boundaries3, puma_boundaries4, puma_boundaries5, puma_boundaries6, puma_boundaries7, puma_boundaries8)

puma_boundaries <- puma_boundaries0

length(unique( puma_data$PUMA))

# add characteristics of each PUMA to be displayed in the table

 puma_data<-import("C:/Users/Zahir Nikmal/Downloads/puma_data.xlsx")

puma_boundaries0<-puma_boundaries0%>%
  left_join(puma_data,by=c("PUMACE10"="PUMA"))%>%
  
  mutate(NAMELSAD10=gsub(" PUMA","",NAMELSAD10))


## Export puma data
puma_boundaries_export<-puma_boundaries0%>%
  select(PUMACE10,NAMELSAD10,contains("perc"),contains("pop20"),contains("abs"),contains("_ov"))%>%st_drop_geometry()
colnames(puma_boundaries_export)<-c("PUMA ID",
"PUMA name","Meat Consumption %", "Milk&Dairy Consumption %", "Fast Food Consumption %", "Fruit Consumption %", "Vegetables Consumption %", "Sugar Consumption %", "Eaten out side %", "Population 20+")


export(puma_boundaries_export,"puma_boundaries_export.xlsx",overwrite=TRUE)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(leaflet)
library(stringr)
library(sf)

# Create color palettes for segment1_perc

palette1 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Meat_perc)

palette2 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Milk_dairy_perc)

palette3 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Fast_food_perc)

palette4 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Fruits_perc)

palette5 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$vegetables_perc)

palette6 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Sugar_perc)

palette7 <- colorNumeric(palette = "viridis", domain = puma_boundaries0$Eat_outside_perc)

# Create the leaflet map
puma_boundaries0 %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Positron") %>%
    addPolygons(
              popup=~paste0(NAMELSAD10,"<br>",Meat_perc),
              fillColor = ~palette1(Meat_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7,
              group="Meat Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",Milk_dairy_perc),
              fillColor = ~palette2(Milk_dairy_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Milk&Dairy Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",Fast_food_perc),
              fillColor = ~palette3(Fast_food_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Fast Food Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",Fruits_perc),
              fillColor = ~palette4(Fruits_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Fruit Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",vegetables_perc),
              fillColor = ~palette5(vegetables_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Vegetables Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",Sugar_perc),
              fillColor = ~palette6(Sugar_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Sugar Consumption %") %>% 
  addPolygons(popup=~paste0(NAMELSAD10,"<br>",Eat_outside_perc),
              fillColor = ~palette7(Eat_outside_perc), 
              weight = 1, 
              opacity = 1, 
              color = "white", 
              fillOpacity = 0.7, 
              group = "Eaten out side %") %>% 
  addLegend("bottomright", pal = palette1, values = ~Meat_perc, title = "Meat Consumption %", opacity = 1, group="Meat Consumption %", layerId="Meat Consumption %") %>% 
  addLegend("bottomright", pal = palette2, values = ~Milk_dairy_perc, title = "Milk&Dairy Consumption %", opacity = 1, group="Milk&Dairy Consumption %", layerId="Milk&Dairy Consumption %") %>%
  addLegend("bottomright", pal = palette3, values = ~Fast_food_perc, title = "Fast Food Consumption %", opacity = 1, group="Fast Food Consumption %", layerId="Fast Food Consumption %") %>%
addLegend("bottomright", pal = palette4, values = ~Fruits_perc, title = "Fruit Consumption %", opacity = 1, group="Fruit Consumption %", layerId="Fruit Consumption %") %>%
addLegend("bottomright", pal = palette5, values = ~vegetables_perc, title = "Vegetables Consumption %", opacity = 1, group="Vegetables Consumption %", layerId="Vegetables Consumption %") %>%
addLegend("bottomright", pal = palette6, values = ~Sugar_perc, title = "Sugar Consumption %", opacity = 1, group="Sugar Consumption %", layerId="Sugar Consumption %") %>%
addLegend("bottomright", pal = palette7, values = ~Eat_outside_perc, title = "Eaten out side %", opacity = 1, group="Eaten out side %", layerId="Eaten out side %") %>%
  addLayersControl(baseGroups = c("Meat Consumption %","Milk&Dairy Consumption %","Fast Food Consumption %","Fruit Consumption %","Vegetables Consumption %","Sugar Consumption %","Eaten out side %"),
                   options = layersControlOptions(collapsed = FALSE))%>% 
hideGroup("Milk&Dairy Consumption %")%>%
hideGroup("Fast Food Consumption %")%>%
hideGroup("Fruit Consumption %")%>%
hideGroup("Vegetables Consumption %")%>%
hideGroup("Sugar Consumption %")%>%
hideGroup("Eaten out side %")%>%
  htmlwidgets::onRender("
    function(el, x) {
      var initialLegend = 'Meat Consumption %' 
      var myMap = this;
      for (var legend in myMap.controls._controlsById) {
        var el = myMap.controls.get(legend.toString())._container;
        if(legend.toString() === initialLegend) {
          el.style.display = 'block';
        } else {
          el.style.display = 'none';
        };
      };
    myMap.on('baselayerchange',
      function (layer) {
        for (var legend in myMap.controls._controlsById) {
          var el = myMap.controls.get(legend.toString())._container;
          if(legend.toString() === layer.name) {
            el.style.display = 'block';
          } else {
            el.style.display = 'none';
          };
        };
      });
    }")

```


## Varible importance for Meat
```{r}
library(dplyr)
vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train <- 
  survey_meat %>% 
  select(all_of(vip_features))


explainer_rf <- 
  explain_tidymodels(
    rf_fit, 
    data = vip_train, 
    y = survey_meat$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex <- survey_meat[120,]

rf_breakdown <- predict_parts(explainer = explainer_rf, new_observation = duplex)

####
library(forcats)

big_house <- vip_train[1269,]

set.seed(1802)
shap_house <- 
  predict_parts(
    explainer = explainer_rf, 
    new_observation = big_house, 
    type = "shap",
    B = 20
  )

shap_house %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

#####

ggplot_pdp <- function(obj, x) {
  
  p <- 
    as_tibble(obj$agr_profiles) %>%
    mutate(`_label_` = stringr::str_remove(`_label_`, "^[^_]*_")) %>%
    ggplot(aes(`_x_`, `_yhat_`)) +
    geom_line(data = as_tibble(obj$cp_profiles),
              aes(x = {{ x }}, group = `_ids_`),
              linewidth = 0.5, alpha = 0.05, color = "gray50")
  
  num_colors <- n_distinct(obj$agr_profiles$`_label_`)
  
  if (num_colors > 1) {
    p <- p + geom_line(aes(color = `_label_`), linewidth = 1.2, alpha = 0.8)
  } else {
    p <- p + geom_line(color = "midnightblue", linewidth = 1.2, alpha = 0.8)
  }
  
  p
}



#######

set.seed(1806)
pdp_liv <- model_profile(explainer_rf, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Grams", 
       color = NULL)

```


## Varible importance for milk

```{r}

vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train1 <- 
  survey_Milk_Dairy %>% 
  select(all_of(vip_features))


explainer_rf1 <- 
  explain_tidymodels(
    rf_fit1, 
    data = vip_train1, 
    y = survey_Milk_Dairy$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex1 <- survey_Milk_Dairy[120,]

rf_breakdown1 <- predict_parts(explainer = explainer_rf1, new_observation = duplex1)

####

big_house1 <- vip_train1[1269,]

set.seed(1802)
shap_house1 <- 
  predict_parts(
    explainer = explainer_rf1, 
    new_observation = big_house1, 
    type = "shap",
    B = 20
  )

shap_house1 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf1, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Dairy by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf1, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Dairy by Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf1, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = " Dairy by Grams", 
       color = NULL)

```

# Varible importance for Fast food

```{r}

vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train2 <- 
  survey_Fast_food %>% 
  select(all_of(vip_features))


explainer_rf2 <- 
  explain_tidymodels(
    rf_fit2, 
    data = vip_train2, 
    y = survey_Milk_Dairy$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex2 <- survey_Fast_food[120,]

rf_breakdown2 <- predict_parts(explainer = explainer_rf2, new_observation = duplex2)

####

big_house2 <- vip_train2[1269,]

set.seed(1802)
shap_house2 <- 
  predict_parts(
    explainer = explainer_rf2, 
    new_observation = big_house2, 
    type = "shap",
    B = 20
  )

shap_house2 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf2, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fast Food by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf2, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fast Food by Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf2, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fast Food by Grams", 
       color = NULL)

```

# Varible importance for Fruit

```{r}
vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train3 <- 
  survey_Fast_food %>% 
  filter(Grams < 250) %>% 
  select(all_of(vip_features))


explainer_rf3 <- 
  explain_tidymodels(
    rf_fit3, 
    data = vip_train3, 
    y = survey_Fruits$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex3 <- survey_Fruits[120,]

rf_breakdown3 <- predict_parts(explainer = explainer_rf3, new_observation = duplex3)

####

big_house3 <- vip_train3[1269,]

set.seed(1802)
shap_house3 <- 
  predict_parts(
    explainer = explainer_rf3, 
    new_observation = big_house3, 
    type = "shap",
    B = 20
  )

shap_house3 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf3, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fruit by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf3, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fruit by Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf3, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Fruit by Grams", 
       color = NULL)

```

# Varible importance for Vegetable

```{r}
vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train4 <- 
  survey_Vegetables %>% 
  filter(Grams < 150 )  %>% 
  select(all_of(vip_features))


explainer_rf4 <- 
  explain_tidymodels(
    rf_fit4, 
    data = vip_train4, 
    y = survey_Vegetables$Grams,
    label = "random forest",
    verbose = FALSE
  )

duplex4 <- survey_Vegetables[120,]

rf_breakdown4 <- predict_parts(explainer = explainer_rf4, new_observation = duplex4)

####

big_house4 <- vip_train4[1269,]

set.seed(1802)
shap_house4 <- 
  predict_parts(
    explainer = explainer_rf4, 
    new_observation = big_house4, 
    type = "shap",
    B = 20
  )

shap_house4 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf4, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Vegetables by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf4, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Vegetables by Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf4, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Vegetables by Grams", 
       color = NULL)

```

# Varible importance for Sugar

```{r}

vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race", "Country of Birth")

vip_train5 <- 
  survey_final %>% 
  select(all_of(vip_features))


explainer_rf5 <- 
  explain_tidymodels(
    rf_fit5, 
    data = vip_train5, 
    y = survey_final$Sugar,
    label = "random forest",
    verbose = FALSE
  )

duplex5 <- survey_final[120,]

rf_breakdown5 <- predict_parts(explainer = explainer_rf5, new_observation = duplex5)

####

big_house5 <- vip_train5[1269,]

set.seed(1802)
shap_house5 <- 
  predict_parts(
    explainer = explainer_rf5, 
    new_observation = big_house5, 
    type = "shap",
    B = 20
  )

shap_house5 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf5, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Sugar by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf5, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Sugar by Grams", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf5, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Sugar by Grams", 
       color = NULL)

```

# Varible importance for home out side

```{r}

vip_features <- c("AGE", "EDUCATION", "MARITAL", 
                  "Gender", "Race")

vip_train6 <- 
  survey_final %>% 
  select(all_of(vip_features))


explainer_rf6 <- 
  explain_tidymodels(
    rf_fit6, 
    data = vip_train6, 
    y = survey_final$Did_eat_the_food_at_home,
    label = "random forest",
    verbose = FALSE
  )

duplex6 <- survey_final[120,]

rf_breakdown6 <- predict_parts(explainer = explainer_rf6, new_observation = duplex6)

####

big_house6 <- vip_train6[1269,]

set.seed(1802)
shap_house6 <- 
  predict_parts(
    explainer = explainer_rf6, 
    new_observation = big_house6, 
    type = "shap",
    B = 20
  )

shap_house6 %>%
  group_by(variable) %>%
  mutate(mean_val = mean(contribution)) %>%
  ungroup() %>%
  mutate(variable = fct_reorder(variable, abs(mean_val))) %>%
  ggplot(aes(contribution, variable, fill = mean_val > 0)) +
  geom_col(data = ~distinct(., variable, mean_val), 
           aes(mean_val, variable), 
           alpha = 0.5) +
  geom_boxplot(width = 0.5) +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  labs(y = NULL)

###

set.seed(1806)
pdp_liv <- model_profile(explainer_rf6, N = 1000, 
                         variables = "AGE",
                         groups = "Race")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Sugar by Grams", 
       color = NULL)

# By Gender

set.seed(1806)
pdp_liv <- model_profile(explainer_rf6, N = 1000, 
                         variables = "AGE",
                         groups = "Gender")

ggplot_pdp(pdp_liv, AGE) +
  scale_x_log10() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Eat out side", 
       color = NULL)

# By Education

set.seed(1806)
pdp_liv <- model_profile(explainer_rf6, N = 1000, 
                         variables = "AGE",
                         groups = "EDUCATION")

ggplot_pdp(pdp_liv, AGE) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = "Age", 
       y = "Eat out side", 
       color = NULL)


```

\`\`\`{r}
